---
title: "After The Bell"
author: "Katie Fitzgerald"
date: "`r Sys.Date()`"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number-section: false
    highlight: tango
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
```

# Read the Data

```{r}
atb <- read_csv("clean_data_challenge.csv")

head(atb)
#str(atb)
```

# Filter Data

You might have a variable that either has missing data, or data where the parent skipped the question. Sometimes you might want to remove this missing data from your graph. But, not always! Sometimes the missing data is a part of the story. You will need to decide if the missing data is relevant or can be ignored.

You can use the `filter` function in R to remove responses (rows) that meet a certain criteria for removal

```{r}

#creating a graph of the enjoys variable
#The plot with missing values included
atb %>%
  ggplot(aes(x = enjoys)) +
  geom_bar()

#The plot with missing values removed
atb %>%
  filter(enjoys != "Skip") %>%
  ggplot(aes(x = enjoys)) +
  geom_bar()
```

Alternatively, there will be times when you want to filter to INCLUDE only certain response options, instead of removing options. I could have made the same edit by specifying which response options to keep:

```{r}

#creating a graph of the enjoys variable

#The plot with missing values removed
atb %>%
  filter(enjoys %in% c( "Strongly Disagree", "Agree", "Disagree", "Strongly Agree")) %>% #the order I list the levels does not matter
  ggplot(aes(x = enjoys)) +
  geom_bar()

```

There are many other rules you can use to filter data.

If you know there is a filter you want to make permanent, you can also filter the original dataset and store the revised dataset to use in graphs, instead of having to repeat the filter in multiple chunks of code. For example:

```{r}
#create a new dataset where rows are removed 
atb_noskip <- atb %>%
  filter(enjoys != "Skip")

#now make the plot with my new dataset
atb_noskip %>%
  ggplot(aes(x=enjoys)) +
  geom_bar()
```

# Changing the Level Order for a Categorical Variables

## Manual Order

How can you reorder the levels of a categorical variable? For example, the Enjoy variable is order, and we want the levels to appear in a logical order. Similar to colors, R will default to displaying the levels in alphabetical order. We can use the 'factor' function to change the order that the levels will display.

```{r}

atb_noskip %>%
  mutate(enjoys=factor(enjoys, levels=c("Strongly Disagree", "Disagree", "Agree", "Strongly Agree"))) %>%
  ggplot(aes(x=enjoys)) +
  geom_bar()

```

## Order by Another Variable

Remember our Gestalt Principle, the Law of Simplicity? If our variable is nominal (there is not a natural order), we often want to order bars by their size. We saw extensive examples of this in our in-class R project “Visualizing Amounts and Perception” from Day 8 of class. Here is one example below, but see that R project for many more examples of reordering, combining, and wrangling category levels!

You will need to load the `forcats` library for these `fct` functions to work. The forcats package allows us to work directly with categorical variables in ways that make applying Gestalt Principle straightforward. Learn more about forcats here.

We can use `fct_infreq` to apply the principle of simplicity.

Note that `forcats` also has a function, `fct_reorder`, that allows you to reorder levels by sorting along another variable. For example, this would allow you to reorder boxplots in descending order based on the median values. There was also an example of this in the same R project!

## Faceting with proportions removes color

When faceting and plotting proportions, getting fill to work with the x variable is not straightforward. You cannot just do `fill = xvar`.

```{r}
custom_order <- c("Always", "Sometimes", "Rarely", "Never")


atb %>%
  filter(hmwk_check != "Skip") %>%
  mutate(hmwk_check = factor(hmwk_check, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = hmwk_check, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~grades)
```

# EDA

```{r}
library(ggplot2)

atb %>%
  ggplot(aes(x = enjoys)) +
  geom_bar()

atb %>%
  ggplot(aes(x = grades)) +
  geom_bar()

# atb %>%
#   ggplot(aes(x = absent)) +
#   geom_bar()

# atb %>%
#   ggplot(aes(x = attend_event)) +
#   geom_bar()

atb %>%
  ggplot(aes(x = adult_volunteer)) +
  geom_bar()

atb %>%
  ggplot(aes(x = attend_meeting)) +
  geom_bar()

atb %>%
  ggplot(aes(x = attend_pto)) +
  geom_bar()

atb %>%
  ggplot(aes(x = pt_conf)) +
  geom_bar()

# atb %>%
#   ggplot(aes(x = fundraise)) +
#   geom_bar()

atb %>%
  ggplot(aes(x = serve_committee)) +
  geom_bar()

atb %>%
  ggplot(aes(x = met_counselor)) +
  geom_bar()

atb %>%
  ggplot(aes(x = frq_meet)) +
  geom_histogram()

# atb %>%
#   ggplot(aes(x = notes)) +
#   geom_bar()
# 
# atb %>%
#   ggplot(aes(x = newsletter)) +
#   geom_bar()

# atb %>%
#   ggplot(aes(x = school_satisfaction)) +
#   geom_bar()
# 
# atb %>%
#   ggplot(aes(x = teacher_satisfaction)) +
#   geom_bar()
# 
# atb %>%
#   ggplot(aes(x = standard_satisfaction)) +
#   geom_bar()
# 
# atb %>%
#   ggplot(aes(x = discipline_satisfaction)) +
#   geom_bar()
# 
# atb %>%
#   ggplot(aes(x = staff_satisfaction)) +
#   geom_bar()

# atb %>%
#   ggplot(aes(x = homework_days)) +
#   geom_bar()
# 
# atb %>%
#   ggplot(aes(x = homework_hours)) +
#   geom_histogram()

atb %>%
  ggplot(aes(x = child_hmwk_amount)) +
  geom_bar()

# atb %>%
#   ggplot(aes(x = hmwk_place)) +
#   geom_bar()

atb %>%
  ggplot(aes(x = hmwk_check)) +
  geom_bar()

atb %>%
  ggplot(aes(x = hmwk_help)) +
  geom_bar()

atb %>%
  ggplot(aes(x = story)) +
  geom_bar()

atb %>%
  ggplot(aes(x = craft)) +
  geom_bar()

# atb %>%
#   ggplot(aes(x = games)) +
#   geom_bar()

# atb %>%
#   ggplot(aes(x = project)) +
#   geom_bar()

atb %>%
  ggplot(aes(x = sports)) +
  geom_bar()

atb %>%
  ggplot(aes(x = time_manage)) +
  geom_bar()
# 
# atb %>%
#   ggplot(aes(x = fam_history)) +
#   geom_bar()

# atb %>%
#   ggplot(aes(x = eat_together)) +
#   geom_bar()

atb %>%
  ggplot(aes(x = library)) +
  geom_bar()

atb %>%
  ggplot(aes(x = bookstore)) +
  geom_bar()

# atb %>%
#   ggplot(aes(x = household_size)) +
#   geom_bar()
# 
# atb %>%
#   ggplot(aes(x = first_guardian_work)) +
#   geom_bar()
```

notable variables

hmwk_help

library, bookstore, sports, craft, story, time_manage

adult_volunteer

attend_meeting, attend_pto, pt_conf, serve_committee, met_counselor, frq_meet

child_hmwk_amount

hmwk_check

```{r}
### HOMEWORK CHECK
custom_order <- c("Always", "Sometimes", "Rarely", "Never")
atb %>%
  filter(hmwk_check != "Skip") %>%
  mutate(hmwk_check = factor(hmwk_check, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = hmwk_check, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~grades)

atb %>%
  filter(hmwk_check != "Skip") %>%
  mutate(hmwk_check = factor(hmwk_check, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = grades, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~hmwk_check)
### these findings are lowkey strange

## what's up with no grades???

### HOMEWORK HELP
custom_order <- c("Never", "Less than once", "1-2 days", "3-4 days", "5+ days")
atb %>%
  filter(hmwk_help != "Skip") %>%
  mutate(hmwk_help = factor(hmwk_help, levels = custom_order
                             )) %>%
  ggplot() +
  geom_bar(mapping = aes(x = hmwk_help, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~grades)

atb %>%
  filter(hmwk_help != "Skip") %>%
  mutate(hmwk_help = factor(hmwk_help, levels = custom_order
                             )) %>%
  ggplot() +
  geom_bar(mapping = aes(x = grades, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~hmwk_help)
# what's up with never??? is it just the smart kids? it looks like grades go down as homework help goes up...??
```

```{r}
#regression with grades
# 
# library(nnet)
# 
# # Example dataset
# atb <- data.frame(
#   grades = factor(sample(c("A", "B", "C", "D or F", "No grades", "Skip"), 100, replace = TRUE)), # Multiclass response
#   hmwk_help = factor(sample(c("Never", "Less than once", "1-2 days", "3-4 days", "5+ days"), 100, replace = TRUE)) # Categorical predictor
# )
# 
# # Fit the multinomial logistic regression model
# model <- multinom(grades ~ hmwk_help, data = atb)
# 
# # Model summary
# summary(model)
```

```{r}
### HOMEWORK CHECK & enjoys
custom_order <- c("Always", "Sometimes", "Rarely", "Never")
atb %>%
  filter(hmwk_check != "Skip") %>%
  mutate(hmwk_check = factor(hmwk_check, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = hmwk_check, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~enjoys)

atb %>%
  filter(hmwk_check != "Skip") %>%
  mutate(hmwk_check = factor(hmwk_check, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = enjoys, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~hmwk_check)


### HOMEWORK HELP and enjoys
custom_order <- c("Never", "Less than once", "1-2 days", "3-4 days", "5+ days")
atb %>%
  filter(hmwk_help != "Skip") %>%
  mutate(hmwk_help = factor(hmwk_help, levels = custom_order
                             )) %>%
  ggplot() +
  geom_bar(mapping = aes(x = hmwk_help, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~enjoys)

atb %>%
  filter(hmwk_help != "Skip") %>%
  mutate(hmwk_help = factor(hmwk_help, levels = custom_order
                             )) %>%
  ggplot() +
  geom_bar(mapping = aes(x = enjoys, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~hmwk_help)
##this may actually be saying something... seems like as hw help goes up, there is more agreement about enjoying school, BUT ONLY SLIGHTLY
```

```{r}
#custom_order <- c("Always", "Sometimes", "Rarely", "Never")
atb %>%
  filter(adult_volunteer != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = adult_volunteer, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~enjoys)

atb %>%
  filter(adult_volunteer != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = enjoys, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~adult_volunteer)
### THIS IS WEIRD
```

```{r}
###LIBRARY AND GRADES
atb %>%
  filter(library != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = library, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~grades)

atb %>%
  filter(grades != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = grades, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~library)


###SERVE COMMITTEE AND GRADES
atb %>%
  filter(serve_committee != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = serve_committee, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~grades)

atb %>%
  filter(serve_committee != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = grades, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~serve_committee)

###these are more promising; seems like when they serve on a committee, students tend to have more As

### MET COUNSELOR AND GRADES

atb %>%
  filter(attend_pto != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = attend_pto, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~enjoys)

atb %>%
  filter(attend_pto != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = enjoys, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~attend_pto)
```

```{r}
### SATISFACTION

#ggplot(atb, aes(x=))

atb %>%
  filter(adult_volunteer != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = teacher_satisfaction, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~adult_volunteer)




atb %>%
  filter(enjoys != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = school_satisfaction, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~enjoys)



atb %>%
  filter(time_manage != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = enjoys, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~time_manage)

```

# Visualizations

```{r}

custom_order <- c("Strongly Agree", "Agree", "Disagree", "Strongly Disagree")
# 
# 
# atb %>%
#   filter(hmwk_check != "Skip") %>%
#   mutate(hmwk_check = factor(hmwk_check, levels = custom_order)) %>%
#   ggplot() +
#   geom_bar(mapping = aes(x = hmwk_check, fill = factor(..x..), y = after_stat(prop), group=1)) +
#   scale_fill_viridis_d() +
#   facet_wrap(~grades)



atb %>%
  filter(hmwk_help != "Skip") %>%
  mutate(enjoys = factor(enjoys, levels = custom_order
                             )) %>%
  ggplot() +
  geom_bar(mapping = aes(x = enjoys, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~hmwk_help) +
  labs(y="Proportion",
       x="Parental agreement/disagreement with the statement: \"This child enjoys school.\"",
       ) 


```

```{r}
atb %>%
  filter(serve_committee != "Skip") %>%
  #mutate(adult_volunteer = factor(adult_volunteer, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = grades, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~serve_committee)
```
