---
title: "After The Bell"
author: "Katie Fitzgerald"
date: "`r Sys.Date()`"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number-section: false
    highlight: tango
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
```

# Read the Data

```{r}
atb <- read_csv("clean_data_challenge.csv")

head(atb)
#str(atb)
```

# Filter Data

You might have a variable that either has missing data, or data where the parent skipped the question. Sometimes you might want to remove this missing data from your graph. But, not always! Sometimes the missing data is a part of the story. You will need to decide if the missing data is relevant or can be ignored.

You can use the `filter` function in R to remove responses (rows) that meet a certain criteria for removal

```{r}

#creating a graph of the enjoys variable
#The plot with missing values included
atb %>%
  ggplot(aes(x = enjoys)) +
  geom_bar()

#The plot with missing values removed
atb %>%
  filter(enjoys != "Skip") %>%
  ggplot(aes(x = enjoys)) +
  geom_bar()
```

Alternatively, there will be times when you want to filter to INCLUDE only certain response options, instead of removing options. I could have made the same edit by specifying which response options to keep:

```{r}

#creating a graph of the enjoys variable

#The plot with missing values removed
atb %>%
  filter(enjoys %in% c( "Strongly Disagree", "Agree", "Disagree", "Strongly Agree")) %>% #the order I list the levels does not matter
  ggplot(aes(x = enjoys)) +
  geom_bar()

```

There are many other rules you can use to filter data.

If you know there is a filter you want to make permanent, you can also filter the original dataset and store the revised dataset to use in graphs, instead of having to repeat the filter in multiple chunks of code. For example:

```{r}
#create a new dataset where rows are removed 
atb_noskip <- atb %>%
  filter(enjoys != "Skip")

#now make the plot with my new dataset
atb_noskip %>%
  ggplot(aes(x=enjoys)) +
  geom_bar()
```

# Changing the Level Order for a Categorical Variables

## Manual Order

How can you reorder the levels of a categorical variable? For example, the Enjoy variable is order, and we want the levels to appear in a logical order. Similar to colors, R will default to displaying the levels in alphabetical order. We can use the 'factor' function to change the order that the levels will display.

```{r}

atb_noskip %>%
  mutate(enjoys=factor(enjoys, levels=c("Strongly Disagree", "Disagree", "Agree", "Strongly Agree"))) %>%
  ggplot(aes(x=enjoys)) +
  geom_bar()

```

## Order by Another Variable

Remember our Gestalt Principle, the Law of Simplicity? If our variable is nominal (there is not a natural order), we often want to order bars by their size. We saw extensive examples of this in our in-class R project “Visualizing Amounts and Perception” from Day 8 of class. Here is one example below, but see that R project for many more examples of reordering, combining, and wrangling category levels!

You will need to load the `forcats` library for these `fct` functions to work. The forcats package allows us to work directly with categorical variables in ways that make applying Gestalt Principle straightforward. Learn more about forcats here.

We can use `fct_infreq` to apply the principle of simplicity.

Note that `forcats` also has a function, `fct_reorder`, that allows you to reorder levels by sorting along another variable. For example, this would allow you to reorder boxplots in descending order based on the median values. There was also an example of this in the same R project!

## Faceting with proportions removes color

When faceting and plotting proportions, getting fill to work with the x variable is not straightforward. You cannot just do `fill = xvar`.

```{r}
custom_order <- c("Always", "Sometimes", "Rarely", "Never")


atb %>%
  filter(hmwk_check != "Skip") %>%
  mutate(hmwk_check = factor(hmwk_check, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = hmwk_check, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~grades)
```

```{r}
```

```{r}
#star



atb %>%
  filter(total_income != "Skip") %>%
  #mutate(first_guardian_employment = factor(first_guardian_employment, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = total_income, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~adult_volunteer)
```

```{r}
#star star


atb %>%
  filter(first_guardian_marital != "Skip") %>%
  #mutate(first_guardian_employment = factor(first_guardian_employment, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = first_guardian_marital, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~adult_volunteer)
```

```{r}

#star


atb %>%
  filter(first_guardian_hours != "Skip") %>%
  #mutate(first_guardian_employment = factor(first_guardian_employment, levels = custom_order)) %>%
  ggplot() +
  geom_histogram(mapping = aes(x = first_guardian_hours)) +
  scale_fill_viridis_d() +
  facet_wrap(~adult_volunteer)


```

```{r}

#star star

atb %>%
  filter(first_guardian_hours != "Skip") %>%
  #mutate(first_guardian_employment = factor(first_guardian_employment, levels = custom_order)) %>%
  ggplot() +
  geom_histogram(mapping = aes(x = first_guardian_hours)) +
  scale_fill_viridis_d() +
  facet_wrap(~hmwk_help)
```

```{r}
#star star


atb %>%
  filter(future_ed != "Skip") %>%
  #mutate(first_guardian_employment = factor(first_guardian_employment, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = future_ed, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~attend_pto)
```

```{r}
#star star star

atb %>%
  filter(eat_together != "Skip") %>%
  #mutate(first_guardian_employment = factor(first_guardian_employment, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = eat_together, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~hmwk_help)
```

```{r}
atb %>%
  filter(guardian_type != "Skip") %>%
  #mutate(first_guardian_employment = factor(first_guardian_employment, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = guardian_type, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~hmwk_help)

```

```{r}
#star star star star. Caylee

atb %>%
  filter(total_income != "Skip") %>%
  #mutate(first_guardian_employment = factor(first_guardian_employment, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = total_income, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~adult_volunteer) +
  coord_flip()
```

```{r}
atb %>%
  filter(total_income != "Skip") %>%
  #mutate(first_guardian_employment = factor(first_guardian_employment, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = total_income, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~adult_volunteer) +
  coord_flip()
```

```{r}
atb %>%
  filter(total_income != "Skip") %>%
  #mutate(first_guardian_employment = factor(first_guardian_employment, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = total_income, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~adult_volunteer) +
  coord_flip()
```

```{r}
custom_order <- c("Divorced", "Never married", "Separated", "Widowed", "Now married")

atb %>%
  filter(first_guardian_marital != "Skip") %>%
  filter(adult_volunteer != "Skip") |> 
  mutate(first_guardian_marital = factor(first_guardian_marital, levels = custom_order)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = first_guardian_marital, fill = factor(..x..), y = after_stat(prop), group=1)) +
  scale_fill_viridis_d() +
  facet_wrap(~adult_volunteer) + 
  coord_flip()

atb %>%
  filter(first_guardian_marital != "Skip") %>%
  filter(adult_volunteer != "Skip") |> 
  mutate(first_guardian_marital = factor(first_guardian_marital, levels = custom_order)) %>%
  # group_by(adult_volunteer) |> 
  # count(first_guardian_marital) |> 
  # mutate(prop_marital = n/sum(n)) |> 
  ggplot() +
  geom_bar(aes(y = adult_volunteer, fill = first_guardian_marital), position = "fill") +
  labs(
    x = "Proportion Of Guardians Participating In Adult Volunteering",
    y = "Whether Or Not Guardians Participate In Adult Volunteering",
    fill = "Fist Guardian Marital Status",
    title = "First Gruardians That Are Curently Married Tend To Participate More In Adult Volunteering",
    subtitle = "First Gruardians That Are Divorces Tend To Participate Less In Adult Volunteering"
  ) +
  scale_fill_viridis_d() +
  theme_minimal() 
```

```{r}
#star star star

custom_order <- c("None", "One day", "Two days", "Three days", "Four days", "Five days", "Six days", "Seven days")

custom_order2 <- c("Skip", "Never", "Less than once", "1-2 days", "3-4 days", "5+ days")

facet_label <- c("None" = "Never Eat Together",
                 "One day" = "Eat Together One Day Of The Week",
                 "Two days" = "Eat Together Two Days Of The Week",
                 "Three days" = "Eat Together Three Days Of The Week",
                 "Four days" = "Eat Together Four Days Of The Week",
                 "Five days" = "Eat Together Five Days Of The Week",
                 "Six days" = "Eat Together Six Days Of The Week",
                 "Seven days" = "Eat Together Seven Days Of The Week")

# atb %>%
#   filter(eat_together != "Skip") %>%
#   filter(hmwk_help != "Skip") |> 
#   mutate(eat_together = factor(eat_together, levels = custom_order)) %>%
#   mutate(hmwk_help = factor(hmwk_help, levels = custom_order2)) |> 
#   mutate(seven_days = if_else(eat_together == "Seven days", 1, 0)) |>
#   ggplot() +
#   geom_bar(mapping = aes(x = eat_together, fill = factor(..x..), y = after_stat(prop), 
#                          group=1)) +
#   scale_fill_viridis_d() +
#   facet_wrap(~hmwk_help, ncol = 1) +
#   coord_flip() +
#   labs(
#     y = "Proportion of Guardians Helping Children With Home Work",
#     x = "How Often Families Eat Together",
#     title = "Families That Eat Together 7 Days Have Helped Their Children With Home Works More Than The Rest"
#   ) 
  

atb %>%
  filter(eat_together != "Skip") %>%
  filter(hmwk_help != "Skip") |> 
  mutate(eat_together = factor(eat_together, levels = custom_order)) %>%
  mutate(hmwk_help = factor(hmwk_help, levels = custom_order2)) |> 
  group_by(eat_together) |> 
  count(hmwk_help) |> 
  mutate(prop_hmwk_help = n/sum(n)) |> 
  mutate(never = case_when(
    hmwk_help == "Never" ~ "yes",
    hmwk_help == "5+ days" ~ "yesyes",
    TRUE ~ "no"
  )) |> 
  ggplot() +
  geom_col(mapping = aes(x = prop_hmwk_help, y = hmwk_help, fill = never)) +
  scale_fill_viridis_d() +
  facet_wrap(~eat_together, labeller = labeller(eat_together = facet_label), ncol = 1) +
  labs(
     x = "Proportion of Guardians Helping Children With Home Work",
     y = "How Often Families Help Their Children With Home Work",
     title = "Families That Eat Together 7 Days Have Helped Their Children With Home Works More Than The Rest"
   ) +
  theme(legend.position = "none")
```

```{r}
library(ggplot2)
library(dplyr)

custom_order <- c("None", "One day", "Two days", "Three days", "Four days", "Five days", "Six days", "Seven days")
custom_order2 <- c("Skip", "Never", "Less than once", "1-2 days", "3-4 days", "5+ days")

# Modify the code to highlight a specific bar (e.g., "Seven days")
atb %>%
  filter(eat_together != "Skip") %>%
  filter(hmwk_help != "Skip") %>% 
  mutate(eat_together = factor(eat_together, levels = custom_order)) %>%
  mutate(hmwk_help = factor(hmwk_help, levels = custom_order2)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = eat_together, 
                         fill = factor(ifelse(eat_together == "Seven days", "highlight", "other")), 
                         y = after_stat(prop), group = 1)) +
  scale_fill_manual(values = c("highlight" = "red", "other" = "grey")) +  # Highlight in red, others in grey
  facet_wrap(~hmwk_help) +
  coord_flip() +
  labs(
    y = "Proportion of Guardians Helping Children With Home Work",
    x = "How Often Families Eat Together",
    title = "Families That Eat Together 7 Days Have Helped Their Children With Home Works More Than The Rest"
  )

```
